---
title: "Assignment4"
author: "Anthony DiStefano"
date: "October 4, 2018"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




#1 Write a function: input a number x and return x^2+3x+1
```{r}
cubicf = function(x)
 {
    return(x^2 + 3*x + 1) 
    }
 cubicf(3)
```


#2	Write a function: input a vector and return the product of all the entries in the vector
```{r}
vector1 <- function(x)
 {
     return(prod(x))
 }
 vector1(c(1,2,3,4))

```

#3 Write a function: input a vector v and a number n and return a vector with its missing data being replaced by n
```{r}
vector2<- function(v,n)
 {
     v[is.na(v)]<- n
     return(v)
 }
v5<- c(5,NA,6)
vector2(v5,6)
```

#4	Write a function: input a numeric vector and return a vector with its missing data being replaced by the mean of the vector
```{r}
 F4<- function(x)
 {
     avg <- mean(x,na.rm = TRUE)
     x[is.na(x)] <- avg
     return(x)
 }
 F4(c(NA,2,3,4))
```

#5 Write a function: input a data frame and return a data frame with numeric missing values being replaced by the mean of the corresponding column.  Check your function on the titanic data
```{r}
F5<- function(df) {
  varcount<- ncol(df)
  for(i in 1:varcount) {
    if(is.numeric(df[[i]]) == TRUE) {
      df[[i]][is.na(df[[i]])] <- mean(df[[i]], na.rm = TRUE)
    }
  }
  return(df)
}

```


#6	Write a function that plots a bar chart of an input categorical variable.  Check your function on the titanic data
```{r}
Num6 <- function(df, Var1) {
    library(ggplot2)
     rt = ggplot(data = df) + geom_bar(mapping = aes(x = df[,Var1]))
     return(rt)
}
Num6(Titanic,2)
```

#7	Write a function that plots bar chart of two input categorical variables.  Check your function on the titanic data
```{r}
F7 <- function(df, variable1, variable2) {
    library(ggplot2)
     rt = ggplot(data = df) + geom_bar(mapping = aes(x = df[,variable1], fill = df[,variable2]), position = "dodge")
     return(rt)
 }
 F7(Titanic, 4,2)
```
